\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2024-03-20 RefereeResponse/paper/2loopNLSM.tex
Redline corrections

The color factor
$C_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,Carrasco:2022jxn,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $N_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors. In this 
form the color factors $C_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
C_i+C_j+C_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}. The duality between color and kinematics is the statement that that there exists a way to rearrange factors between
the various kinematic numerators $N_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
N_i +N_j +N_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for four-point NLSM scattering at tree level produces the numerators
\begin{align}\label{pionNums}
&N^{\text{NLSM}}_s = s(t-u)\\
&N^{\text{NLSM}}_t = t(u-s)\\
&N^{\text{NLSM}}_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels.  The net result is
that the amplitude is a local function, just as one would expect for
the pion four-point tree amplitude.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states. Locality is physically motivated
since Feynman rules produce polynomial numerators.  Indeed, the most
natural way to guarantee locality and avoid spurious poles is to
require the off-shell numerators to be polynomial functions of
kinematics.  However, insisting on a local numerator is partly a
matter of convenience as the ansatz for a rational function would
quickly grow out of control without a guiding principle for what terms
to include.  %Referee edit
Historically, locality has also proved a highly successful tool for
simplifying kinematic
numerators.
For example, the four-point function for $\mathcal{N}=4$ super-Yang
Mills admits local color-dual numerators through four loops
\cite{Bern:2012uf}.\footnote{At five loops, the most general local
  ansatz has not been explored, but the double copy has been performed
  at five loops using alternative methods \cite{GeneralizedDoubleCopy,
    GeneralizedDoubleCopyFiveLoops}.}  While locality is
well motivated from a physical and complexity standpoint, the
literature has featured numerators that violate locality, modify naive
power counting through the inclusion of higher spin modes, and break
manifest Lorentz invariance \cite{Square, WeinzierlBCJLagrangian,
  Mogull:2015adi, FivePointN4BCJ, Johansson:2017bfl,
  Ben-Shahar:2022ixa, Cheung:2016prv, Cheung:2021zvb}.  These
developments suggest that relaxing locality is a viable alternative
when the condition proves too stringent for an ansatz.  However,
simply requiring on-shell locality may suffice.  This alternative will
be discussed in \cref{nonLocalScattering}.

@@

At this point color-kinematics duality has been satisfied, so the
double copy will proceed without issue.  However, the plethora of
generalized gauge freedom parameters leaves the option for enforcing
additional aesthetic constraints.  All remaining generalized gauge
parameters could be set to zero, but a simpler and more insightful
result can be obtained through physical arguments.  $\mathcal{N}=4$
SYM provides several hints for further conditions to impose on the
ansatz.  For example, for maximally supersymmetric gauge theory it is
possible to enforce the no triangle hypothesis, manifest loop power
counting, and, for four-point and up to at least six loops, strip off
a factor of $st \atree$ from the integrand \cite{FiveLoopN4,
  Bern:2012uf, Bern:2010fy, SuperSum, Carrasco:2021otn,
  JJHenrikReview, BRY, BDDPR, Neq44np}.  For the NLSM integrand it is
desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the one-loop numerator
since this at least manifests antisymmetry for the vertices involving
external legs.  However, a more fruitful direction is to match onto
the one known theory with pion power counting that manifests
color-kinematics duality to all loop orders, namely,
Zakharov-Mikhailov (ZM) theory \cite{Zakharov:1973pp, Cheung:2022mix}.
ZM theory is governed by the Lagrangian
\begin{equation}
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2
+ g f^{abc} \varphi^a \varepsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c),
\end{equation}
where more details can be found in \cref{sec:CKLagrangians}.  The
color-stripped Feynman rule for the vertex is
$V(k_1, k_2, k_3) \propto \varepsilon_{\mu\nu}k_1^\mu k_2^\nu \equiv \langle k_1 k_2 \rangle$ where
off-shell color-kinematics duality to all orders in perturbation
theory simply follows from the Schouten identity in 2D.  Since the theory
is purely cubic and manifests off-shell color-kinematics
duality, it is trivial to read off the color-dual numerator for any
graph.  From the presence of the Levi-Civita tensor
$\varepsilon^{\mu\nu}$, the theory clearly resides in two spacetime
dimensions where scattering is notoriously plagued by infrared
regulation issues.  Every on-shell particle is either a right mover,
with momentum proportional to $k_R^\mu \equiv (1,1)$, or a left mover,
with momentum proportional to $k_L^\mu \equiv (1,-1)$.  On-shell ZM
amplitudes naturally divide into sectors corresponding to the
configuration of left and right movers, where the scattering in many
sectors is rather subtle.  Only color-ordered amplitudes can be
defined unambiguously from the naive Feynman rules.  At four-point, only the
alternating sector is free of subtleties and the amplitude for this
process vanishes.  In equations, $A[LRLR]=0$ where $R$ corresponds to
a right mover and $L$ corresponds to a left mover.  In every other
configuration of right and left movers, such as $A[RRRR]$, one of the
internal propagators is accidentally on-shell since $k_R^2=k_L^2=0$.

@@

To construct the double-copy theories, we source the cubic
pure YM numerators from Ref.~\cite{Bern:2015ooa} (which
additionally satisfies a relaxed form of color-kinematics duality),
and the cubic $\mathcal{N}=4$ sYM representation from the well-known
$n_{2\text{box}} = n_{\text{cross-box}} = s^2t\, \atree$ \cite{Bern:1997nh}.
On the other hand, we use the methods of Ref.~\cite{Carrasco:2023qgz}
to directly compute the needed basis cuts in all three theories without
relying on a color-dual pion representation.  We find exact agreement
in each of the three theories for both physical cuts.  Since
Ref.~\cite{Carrasco:2023qgz} has already exhaustively explored the
properties of four-point loop amplitudes in these theories, we direct
interested readers there for more information.

>>>>

The color factor
$C_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,Carrasco:2022jxn,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $N_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors. In this 
form the color factors $C_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
C_i+C_j+C_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}. The duality between color and kinematics is the statement that there exists a way to rearrange factors between
the various kinematic numerators $N_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
N_i +N_j +N_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for four-point NLSM scattering at tree level produces the numerators
\begin{align}\label{pionNums}
&N^{\text{NLSM}}_s = s(t-u)\\
&N^{\text{NLSM}}_t = t(u-s)\\
&N^{\text{NLSM}}_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels.  The net result is
that the amplitude is a local function, just as one would expect for
the pion four-point tree amplitude.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states. Locality is physically motivated
since Feynman rules produce polynomial numerators.  Indeed, the most
natural way to guarantee locality and avoid spurious poles is to
require the off-shell numerators to be polynomial functions of
kinematics.  However, insisting on a local numerator is partly a
matter of convenience as the ansatz for a rational function would
quickly grow out of control without a guiding principle for what terms
to include.  %Referee edit
Historically, locality has also proved a highly successful tool for
simplifying kinematic
numerators.
For example, the four-point function for $\mathcal{N}=4$ super-Yang
Mills admits local color-dual numerators through four loops
\cite{Bern:2012uf}.\footnote{At five loops, the most general local
  ansatz has not been explored, but the double copy has been performed
  at five loops using alternative methods \cite{GeneralizedDoubleCopy,
    GeneralizedDoubleCopyFiveLoops}.}  While locality is
well motivated from a physical and complexity standpoint, the
literature has featured numerators that violate locality, modify naive
power counting through the inclusion of higher spin modes, and break
manifest Lorentz invariance \cite{KiermaierTalk, WeinzierlBCJLagrangian,
  Mogull:2015adi, FivePointN4BCJ, Johansson:2017bfl,
  Ben-Shahar:2022ixa, Cheung:2016prv, Cheung:2021zvb}.  These
developments suggest that relaxing locality is a viable alternative
when the condition proves too stringent for an ansatz.  However,
simply requiring on-shell locality may suffice.  This alternative will
be discussed in \cref{nonLocalScattering}.

@@

At this point color-kinematics duality has been satisfied, so the
double copy will proceed without issue.  However, the plethora of
generalized gauge freedom parameters leaves the option for enforcing
additional aesthetic constraints.  All remaining generalized gauge
parameters could be set to zero, but a simpler and more insightful
result can be obtained through physical arguments.  $\mathcal{N}=4$
SYM provides several hints for further conditions to impose on the
ansatz.  For example, for maximally supersymmetric gauge theory it is
possible to enforce the no triangle hypothesis, manifest loop power
counting, and, for four-point and up to at least six loops, strip off
a factor of $st \atree$ from the integrand \cite{FiveLoopN4,
  Bern:2012uf, Bern:2010fy, SuperSum, Carrasco:2021otn,
  JJHenrikReview, BRY, BDDPR, Neq44np}.  For the NLSM integrand it is
desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the one-loop numerator
since this at least manifests antisymmetry for the vertices involving
external legs.  However, a more fruitful direction is to match onto
the one known theory with pion power counting that manifests
color-kinematics duality to all loop orders, namely,
Zakharov-Mikhailov (ZM) theory \cite{Zakharov:1973pp, Cheung:2022mix}.
ZM theory is governed by the Lagrangian
\begin{equation}
%\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2
+ g f^{abc} \varphi^a \varepsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c),
\end{equation}
where more details can be found in \cref{sec:CKLagrangians}.  The
color-stripped Feynman rule for the vertex is
$V(k_1, k_2, k_3) \propto \varepsilon_{\mu\nu}k_1^\mu k_2^\nu \equiv \langle k_1 k_2 \rangle$ where
off-shell color-kinematics duality to all orders in perturbation
theory simply follows from the Schouten identity in 2D.  Since the theory
is purely cubic and manifests off-shell color-kinematics
duality, it is trivial to read off the color-dual numerator for any
graph.  From the presence of the Levi-Civita tensor
$\varepsilon^{\mu\nu}$, the theory clearly resides in two spacetime
dimensions where scattering is notoriously plagued by infrared
regulation issues.  Every on-shell particle is either a right mover,
with momentum proportional to $k_R^\mu \equiv (1,1)$, or a left mover,
with momentum proportional to $k_L^\mu \equiv (1,-1)$.  On-shell ZM
amplitudes naturally divide into sectors corresponding to the
configuration of left and right movers, where the scattering in many
sectors is rather subtle.  Only color-ordered amplitudes can be
defined unambiguously from the naive Feynman rules.  At four-point, only the
alternating sector is free of subtleties and the amplitude for this
process vanishes.  In equations, $A[LRLR]=0$ where $R$ corresponds to
a right mover and $L$ corresponds to a left mover.  In every other
configuration of right and left movers, such as $A[RRRR]$, one of the
internal propagators is accidentally on-shell since $k_R^2=k_L^2=0$.

@@

To construct the double-copy theories, we source the cubic
pure YM numerators from Ref.~\cite{Bern:2015ooa} (which
additionally satisfies a relaxed form of color-kinematics duality),
and the cubic $\mathcal{N}=4$ sYM representation from the well-known
$n_{2\text{box}} = n_{\text{cross-box}} = s^2t\, \atree$ \cite{BRY}.
On the other hand, we use the methods of Ref.~\cite{Carrasco:2023qgz}
to directly compute the needed basis cuts in all three theories without
relying on a color-dual pion representation.  We find exact agreement
in each of the three theories for both physical cuts.  Since
Ref.~\cite{Carrasco:2023qgz} has already exhaustively explored the
properties of four-point loop amplitudes in these theories, we direct
interested readers there for more information.
